---
title: "Homework Assignment 2"
author: "Reynaldo Perez"
date: "October 16, 2022"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---


```{r setup, echo=FALSE}
library(knitr)
# set global chunk options: images will be 7x5 inches
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(digits = 4)


## indents are for indenting r code as formatted text
## They may need to be adjusted depending on your OS
# if your output looks odd, increase or decrease indent
indent1 = '    '
indent2 = '        '
indent3 = '            '
```

```{r}
library(tidyverse)  # Load tidyverse
library(tidymodels)  # Load tidymodels

abalone <- read.csv("/Users/reynaldoperez/Downloads/homework-2-2/data/abalone.csv")  # Read the data set

names(abalone)  # See the names and number of columns of the data set
```

Q1) Let's add a new variable, named $age$, to the data set.

```{r}
age <- abalone$rings + 1.5  # Calculate age

abalone_new <- cbind(abalone, age)  # Add new variable to the dataset

head(abalone_new)  # Check
```
Now, let us assess the distribution of $age$:

```{r}
hist(abalone_new$age, breaks = "Sturges", main = paste("Distribution of Age"))

```

As one can see, the distribution of $age$ is slightly skewed to the left, with the highest peak at between 10 to ~12 years.


Q2) We will now split the abalone data into a training set and a testing set. We will use stratified sampling.

```{r}
set.seed(1115)

abalone_split <- initial_split(abalone_new, prop = 0.75, strata = age)

abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)

```






Q3) Let us create a recipe for predicting the outcome variable, $age$:

```{r}
simple_abalone_recipe <- recipe(age ~ ., data = abalone_train)

simple_abalone_recipe

```

Now, we will complete the recipe:

```{r}
abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = abalone_train) %>% 
  step_dummy_multi_choice(starts_with("type")) %>%
  prep() %>%
  step_interact(terms = ~type_M:shucked_weight) %>%
  step_interact(terms = ~type_F:shucked_weight) %>%
  step_interact(terms = ~type_I:shucked_weight) %>%
  step_interact(terms = ~longest_shell:diameter) %>%
  step_interact(terms = ~shucked_weight:shell_weight) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())
```

Hence, our recipe is finished. Note that we did not include the $rings$ variable in our recipe. This is because obtaining the number of rings is a very time-consuming task, and the other observed measurements would help predict the age much faster. 


Q4) Now, we will create and store a linear regression object:

```{r}
lm_model <- linear_reg() %>%
  set_engine("lm")
```






Q5) We will now develop an empty workflow, and add the model and recipe we created in the previous questions:

```{r}
lm_wflow <- workflow() %>%
  add_model(lm_model) %>%
  add_recipe(abalone_recipe)
```




Q6) Let's now use the $fit()$ object to predict the age of a hypothetical female abalone with the given information. 


```{r}
lm_fit <- fit(lm_wflow, abalone_train)

lm_fit %>%
  extract_fit_parsnip() %>%
  tidy()

x0 <- data.frame(type = "type_F", longest_shell = 0.5, diameter = 0.1, height = 0.3, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)  # Construct a data frame of the given information
x0  # Display data frame


## predict.lm(lm_fit, new_data = x0)  # Predicted age, but received error saying model cannot include NA values
```


Q7) Now, we will assess our model's performance.


```{r}
library(yardstick)

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))  # Develop predicted values for age
abalone_train_res %>% 
  head()
```

Now, we will develop the metric sets:

```{r}
abalone_metrics <- metric_set(rmse, rsq, mae)
## abalone_metrics(abalone_train_res, truth = age, estimate = .pred)  # Error saying length of "truth" and "estimate" must match

```

Then, create a tibble of the model's predicted values:

```{r}
abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
abalone_train_res %>% 
  head()


```

As one can see, the predicted value is not that far off the actual value of age. The $R^2$ value we calculated is the percentage amount that the variability observed in $age$ is explained by the regression model. 